Defending Byzantine Attacks in Ensemble Federated Learning: A Reputation-based Phishing Approach

### 2022-02-11 Manuscript_TIFS/manuscript_TIFS.tex
Conlusion and Experiment

%\usepackage{float}
%\usepackage{multirow}
%\usepackage{tabularx}
\usepackage{authblk}
\usepackage{booktabs}
\usepackage{algorithmicx}  
\usepackage{algpseudocode}
\usepackage{soul}
\usepackage{diagbox}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
% updated with editorial comments 8/9/2021

\begin{document}
\title{Defending Byzantine Attacks in Ensemble Federated Learning: A Reputation-based Phishing Approach}
\author{Beibei~Li,~\IEEEmembership{Member,~IEEE,}
        Peiran~Wang,~\IEEEmembership{Student Member,~IEEE,}
        Qinglei~Kong,~\IEEEmembership{Member,~IEEE,}
        Yuan~Zhang,~\IEEEmembership{Member,~IEEE,}
        and~Rongxing~Lu,~\IEEEmembership{Fellow,~IEEE}
\thanks{This paper is an extended version of the paper titled `FLPhish: Reputation-Based Phishing Byzantine Defense in Ensemble Federated Learning', which was published in IEEE ISCC 2021, and awarded `Best Paper'.}
\thanks{B. Li and P. Wang are with the School of Cyber Science and Engineering, Sichuan Universsity, Chengdu, Sichuan, China 610065. Email: libeibei@scu.edu.cn; wangpeiran@stu.scu.edu.cn.}
\thanks{Q. Kong is with the Future Network of Intelligence Institute, The Chinese University of Hong Kong, Shenzhen, China 518172, and also with The University of Science and Technology of China, Hefei, China 230052. Email: kql8904@163.com.}
\thanks{Y. Zhang is with the School of Computer Science and Engineering, University of Electronic Science and Technology of China, Chengdu, China 610054. Email: zy\_loye@126.com.}
\thanks{R. Lu is with the Faculty of Computer Science, University of New Brunswick, Fredericton, NB, Canada E3B 5A3. Email: rlu1@unb.ca.}
}

@@

\section{Conclusion}
The conclusion goes here.

>>>>

%\usepackage{float}
\usepackage{multirow}
%\usepackage{tabularx}
\usepackage{authblk}
\usepackage{booktabs}
\usepackage{algorithmicx}  
\usepackage{algpseudocode}
\usepackage{soul}
\usepackage{diagbox}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
% updated with editorial comments 8/9/2021

\begin{document}

\title{Defending Byzantine Attacks in Ensemble Federated Learning: A Reputation-based Phishing Approach}
\author{Beibei~Li,~\IEEEmembership{Member,~IEEE,}
        Peiran~Wang,~\IEEEmembership{Student Member,~IEEE,}
        Qinglei~Kong,~\IEEEmembership{Member,~IEEE,}
        Yuan~Zhang,~\IEEEmembership{Member,~IEEE,}
        and~Rongxing~Lu,~\IEEEmembership{Fellow,~IEEE}
\thanks{This paper is an extended version of the paper titled `FLPhish: Reputation-Based Phishing Byzantine Defense in Ensemble Federated Learning', which was published in IEEE ISCC 2021, and awarded `Best Paper'.}
\thanks{B. Li and P. Wang are with the School of Cyber Science and Engineering, Sichuan Universsity, Chengdu, Sichuan, China 610065. Email: libeibei@scu.edu.cn; wangpeiran@stu.scu.edu.cn.}
\thanks{Q. Kong is with the Future Network of Intelligence Institute, The Chinese University of Hong Kong, Shenzhen, China 518172, and also with The University of Science and Technology of China, Hefei, China 230052. Email: kql8904@163.com.}
\thanks{Y. Zhang is with the School of Computer Science and Engineering, University of Electronic Science and Technology of China, Chengdu, China 610054. Email: zy\_loye@126.com.}
\thanks{R. Lu is with the Faculty of Computer Science, University of New Brunswick, Fredericton, NB, Canada E3B 5A3. Email: rlu1@unb.ca.}
}

@@

\begin{table}[t]

  \renewcommand\arraystretch{1.5}
  \centering
  \caption{Parameter Settings}
  \resizebox{0.5\textwidth}{!}{
  \label{tab:my-table}
  \begin{tabular}{ll}
    \bottomrule
    \textbf{Parameter}                    & \textbf{Parameter Value}              \\ \hline
    Client Number                         & 50 clients                                \\ \hline
    Sample Total Number                   & 60000 samples                             \\ \hline
    \multirow{2}{*}{Client Sample Number} & 800 training samples                      \\
                                          & 200 test samples                          \\ \hline
    \multirow{2}{*}{Server Sample Number} & 8000 unlabeled samples                    \\
                                          & 2000 labeled samples                      \\ \hline
    Round Number                          & 10 iterations                                 \\ \hline
    \multirow{2}{*}{Round Sample Number}  & 800 unlabeled samples                     \\
                                          & 200 labeled samples                       \\ \hline
    Deep Learning Model                   & Residual Networks                         \\ \hline
    Byzantine Fractions $p$               & 0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9     \\ \hline
    Imbalance Degree $q$                  & 0.1,0.2,0.5,0.6,0.7,0.8               \\ \hline
    Dataset $d$                         & MNIST,Fashion-MNIST,CIFAR-10              \\ \hline
    \end{tabular}
}
\end{table}

\section{Performance Evaluation}
In this section, we experimentally evaluate our FLPhish (noted as FLPhish(weight) in the following parts) against Byzantine attacks (denial-of-service attacks) under different experiment settings. Furthermore, we compare FLPhish(weight)'s performance with FedAvg \cite{ref_01_GoogleFL}, Median \cite{ref_13_defense}, and Trimmed Mean \cite{ref_13_defense}'s performance, and the evaluation results demonstrate that FLPhish(weight) outperforms these schemes in defending against Byzantine attacks. Furthermore, we compare our proposed aggregation rule, FLPhish(weight) with former aggregation rule used in our previous work \cite{li2021flphish}. Our previous work (noted as FLPhish($\tau$) in the following parts) utilized a threshold to identify the Byzantine clients. Here we set the threshold to 0.1 (noted as FLPhish($\tau=0.1$)), 0.2 (noted as FLPhish($\tau=0.2$)) and 0.5 (noted as FLPhish($\tau=0.5$)).

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table}[t]
  \renewcommand\arraystretch{1.5}
  \centering
  \caption{Evaluated Scheme}
  \resizebox{0.5\textwidth}{!}{%
  \begin{tabular}{ll}
  \bottomrule
  \textbf{Scheme}                           & \textbf{Description}                                                     \\ \hline
  FedAvg                           & FedAvg proposed by Google \cite{ref_01_GoogleFL}                                \\ \hline
  Median                           & Median proposed by Yin \textit{et al.} \cite{ref_16_defense}                  \\ \hline
  Trimmed Mean                     & Trimmed Mean proposed by Yin \textit{et al.} \cite{ref_16_defense}            \\ \hline
  FLPhish($\tau=0.1$) & Our conference version with $\tau=0.1$ \cite{li2021flphish} \\ \hline
  FLPhish($\tau=0.2$) & Our conference version with $\tau=0.2$ \cite{li2021flphish} \\ \hline
  FLPhish($\tau=0.5$) & Our conference version with $\tau=0.5$ \cite{li2021flphish} \\ \hline
  FLPhish(weight)                  & The scheme proposed by this work                                          \\ \hline
  \end{tabular}%
  }
  \end{table}

\begin{figure*}[!htp]
  \centering
    \subfloat[\scriptsize MNIST $q=0.1$]{\includegraphics[width=0.245\textwidth]{figures/rep_mnist_imbalance_1.pdf}}
    \subfloat[\scriptsize MNIST $q=0.2$]{\includegraphics[width=0.245\textwidth]{figures/rep_mnist_imbalance_2.pdf}}
    \subfloat[\scriptsize MNIST $q=0.5$]{\includegraphics[width=0.245\textwidth]{figures/rep_mnist_imbalance_5.pdf}}
    \subfloat[\scriptsize MNIST $q=0.8$]{\includegraphics[width=0.245\textwidth]{figures/rep_mnist_imbalance_9.pdf}}

    \subfloat[\scriptsize Fashion-MNIST $q=0.1$]{\includegraphics[width=0.245\textwidth]{figures/rep_fashion_imbalance_1.pdf}}
    \subfloat[\scriptsize Fashion-MNIST $q=0.2$]{\includegraphics[width=0.245\textwidth]{figures/rep_fashion_imbalance_2.pdf}}
    \subfloat[\scriptsize Fashion-MNIST $q=0.5$]{\includegraphics[width=0.245\textwidth]{figures/rep_fashion_imbalance_5.pdf}}
    \subfloat[\scriptsize Fashion-MNIST $q=0.8$]{\includegraphics[width=0.245\textwidth]{figures/rep_fashion_imbalance_9.pdf}}

    \subfloat[\scriptsize CIFAR-10 {$q=0.1$}]{\includegraphics[width=0.245\textwidth]{figures/rep_cifar_imbalance_1.pdf}}
    \subfloat[\scriptsize CIFAR-10 {$q=0.2$}]{\includegraphics[width=0.245\textwidth]{figures/rep_cifar_imbalance_2.pdf}}
    \subfloat[\scriptsize CIFAR-10 {$q=0.5$}]{\includegraphics[width=0.245\textwidth]{figures/rep_cifar_imbalance_5.pdf}}
    \subfloat[\scriptsize CIFAR-10 {$q=0.8$}]{\includegraphics[width=0.245\textwidth]{figures/rep_cifar_imbalance_9.pdf}}
  \\
  \caption{Reputation values of experiments on MNIST on different imbalance degrees.}
  \label{fig_exp_reputation}
  \vspace{0.2in}
\end{figure*}

\subsection{Experiment Setup}

  \subsubsection{{\textit{The fraction {$p$} of Byzantine clients}}} We evaluate our FLPhish(weight) under the settings of different fractions $p$ of Byzantine clients: 0 (no Byzantine clients), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9.
  \subsubsection{{\textit{The imbalance degree {$q$} of the data}}} We distribute the data in a dataset among all the clients, based on previous study \cite{ref_06_model}. If a dataset consists if $M$ classes of data, the FL clients will be divided into $M$ groups. Data is delivered to a client $c$ in group $m$, with data $m$ accounting for more than $q$ percent. Data is uniformly disseminated among all clients within the same group. The distribution inference of clients' local training data is controlled by the parameter $q$. If $q=\frac{1}{M}$, the clients' local training data are independent and identically distributed (IID). We evaluate our FLPhish(weight) on 6 different $q$ settings: 0.1 (IID), 0.2, 0.5, 0.6, 0.7, and 0.8.
  \subsubsection{{\textit{The number of clients}}} The number of clients is set to be 50 in our experiment.
  \subsubsection{{\textit{The local CNN model used by clients}}} ResNet is employed to perform deep learning tasks in our local client.
  \subsubsection{{\textit{The datasets {$d$}}}} We take three different datasets as our experiment datasets: \begin{itemize}
      \item MNIST: MNIST is a 10-class digit image classification dataset consisting of 60,000 training samples and 10,000 testing samples.
      \item Fashion-MNIST: Fashion-MNIST is a 10-class fashion image classification dataset. It has a predefined training set of 60,000 fashion images and a testing set of 10,000 fashion images.
      \item CIFAR-10: CIFAR-10 is a color image classification dataset. It has predefined 50,000 training examples and 10,000 testing examples. Each example belongs to one of the 10 classes.
  \end{itemize}

  \begin{figure*}[!htp]
    \centering
    \subfloat[\scriptsize Byzantine Fraction 0.9 \& Imbalance Degree 0.2 \& Random Attack]{\includegraphics[width=0.48\textwidth]{figures/bar_byzantines_random_bar.pdf}}
    \subfloat[\scriptsize Byzantine Fraction 0.9 \& Imbalance Degree 0.2 \& Untargeted Attack]{\includegraphics[width=0.48\textwidth]{figures/bar_byzantines_untargeted_bar.pdf}}
    \hfill
    \subfloat[\scriptsize Byzantine Fraction 0.5 \& Imbalance Degree 0.8 \& Random Attack]{\includegraphics[width=0.48\textwidth]{figures/bar_imbalances_random_bar.pdf}}
    \subfloat[\scriptsize Byzantine Fraction 0.5 \& Imbalance Degree 0.8 \& Untargeted Attack]{\includegraphics[width=0.48\textwidth]{figures/bar_imbalances_untargeted_bar.pdf}}
    \caption{Experiments results under different Byzantine fractions and different imbalance degrees.}
    \label{fig_exp_extreme}
  \end{figure*}

  \begin{figure*}[!htp]
    \centering
    \subfloat[\scriptsize MNIST]{\includegraphics[width=0.325\textwidth]{figures/table_random_imbalances_MNIST.pdf}}
    \subfloat[\scriptsize Fashion-MNIST]{\includegraphics[width=0.325\textwidth]{figures/table_random_imbalances_Fashion-MNIST.pdf}}
    \subfloat[\scriptsize CIFAR-10]{\includegraphics[width=0.325\textwidth]{figures/table_random_imbalances_Cifar-10.pdf}}
    \\
    \caption{Accuracy values on Different Imbalance Degrees under Random Attack.}
    \label{fig_table_imbalances_random}
  \end{figure*}

  \begin{figure*}[!htp]
    \centering
    \subfloat[\scriptsize MNIST]{\includegraphics[width=0.325\textwidth]{figures/table_random_byzantines_MNIST.pdf}}
    \subfloat[\scriptsize Fashion-MNIST]{\includegraphics[width=0.325\textwidth]{figures/table_random_byzantines_Fashion-MNIST.pdf}}
    \subfloat[\scriptsize CIFAR-10]{\includegraphics[width=0.325\textwidth]{figures/table_random_byzantines_Cifar-10.pdf}}
    \\
    \caption{Accuracy values on Different Byzantine Fractions under Random Attack.}
    \label{fig_table_byzantines_random}
  \end{figure*}

  \par Each client has 1,000 samples taken from the training dataset. Among the samples, 800 of them are used as training datasets, while another 200 are taken as test datasets. And the server has 10000 testing examples from the testing dataset. 
  \subsubsection{\textit{Evaluated Byzantine Attacks}} We evaluate our FLPhish(weight) against two Byzantine denial-of-service attacks:
  \begin{itemize}
    \item Untargeted Byzantine Attacks: For each Byzantine client, it mislabels the data $l$ to $(l-1)\ mod\ M$ to launch the attacks against FLPhish(weight). The attack is known as the Label-flipping attack. 
    \item Random Byzantine Attacks: For each Byzantine client, it mislabels the labels by returning a randomly chosen result.
    \end{itemize}

  \subsubsection{{\textit{Experiment Environment}}} We conduct all the experiments on a laptop with Intel(R) Core(TM) i7-11800H CPU 2.30GHz and an NVIDIA GeForce RTX 3060 GPU with the video memory of 6 GB. We implement all deep learning models using Keras\footnote{https://keras.io/}.

  \begin{figure*}[!htp]
    \centering
    \subfloat[\scriptsize MNIST]{\includegraphics[width=0.325\textwidth]{figures/table_untargeted_imbalances_MNIST.pdf}}
    \subfloat[\scriptsize Fashion-MNIST]{\includegraphics[width=0.325\textwidth]{figures/table_untargeted_imbalances_Fashion-MNIST.pdf}}
    \subfloat[\scriptsize CIFAR-10]{\includegraphics[width=0.325\textwidth]{figures/table_untargeted_imbalances_Cifar-10.pdf}}
    \\
    \caption{Accuracy values on Different Imbalance Degrees under Untargeted Attack.}
    \label{fig_table_imbalances_untargeted}
  \end{figure*}

  \begin{figure*}[!htp]
    \centering
    \subfloat[\scriptsize MNIST]{\includegraphics[width=0.325\textwidth]{figures/table_untargeted_byzantines_MNIST.pdf}}
    \subfloat[\scriptsize Fashion-MNIST]{\includegraphics[width=0.325\textwidth]{figures/table_untargeted_byzantines_Fashion-MNIST.pdf}}
    \subfloat[\scriptsize CIFAR-10]{\includegraphics[width=0.325\textwidth]{figures/table_untargeted_byzantines_Cifar-10.pdf}}
    \\
    \caption{Accuracy values on Different Byzantine Fractions under Untargeted Attack.}
    \label{fig_table_byzantines_untargeted}
  \end{figure*}

  \subsection{Performance Comparison under Random Attacks}
  From Fig. \ref{fig_exp_extreme}-\ref{fig_table_byzantines_random}, we can see that random Byzantine attacks have poor performance against FLPhish(weight) and FLPhish($\tau$=0.1, 0.2). We refer from the result that a random Byzantine attack has a drawback that it can not gather its influence at one data point as untargeted Byzantine attackers do. Meanwhile, we can see that it has good results in attacking FLPhish($\tau$=0.5) in the CIFAR-10 dataset. From Fig. \ref{fig_exp_reputation}, we can see that the reputation of benign clients in CIFAR-10 experiments rapidly falls under the threshold of 0.5 due to the complexity of the tasks. With the increase of imbalance degree value $q$, the reputation of the benign clients falls more rapidly in the experiments of the CIFAR-10 dataset than in the experiments of the MNIST and the Fashion-MNIST dataset. As benign clients' reputations in the experiments of the CIFAR-10 dataset become under the threshold $\tau$ of 0.5, they are identified as malicious Byzantine clients as well. As many benign clients are falsely identified as Byzantine attackers, the FL server lacks enough trusted FL clients to assist the aggregation, therefore causing a bad performance of FLPhish($\tau$=0.5). Meanwhile, we can see that FLPhish($\tau$=0.1), FLPhish($\tau$=0.2), and FLPhish(weight) stood still against the random Byzantine attacks. The results indicate that the value of the threshold $\tau$ should be considered carefully in the application in some complex tasks.

  \subsection{{Performance Comparison under Untargeted Attacks}} We further evaluate FLPhish(weight) towards untargeted Byzantine attacks in EFL.

    \subsubsection{\ul{Performance Comparison with Different Distributions}} We test our FLPhish(weight) in an experiment environment with a fixed Byzantine client fraction of 0.5 and a variable distribution imbalance value $q$ across tests. The experiment results are shown in Fig. \ref{fig_table_imbalances_untargeted}. We can see that both FLPhish($\tau$=0.1,0.2,0.5) and FLPhish(weight) outperform other schemes. When the imbalance degree $q$ reaches 0.8, the distribution of data becomes very non-IID. In this situation, the FLPhish($\tau$=0.5)'s performance progressively deteriorates. When confronted with a distribution with an imbalance degree of $q=0.8$, each client performs poorly, resulting in a loss of reputation. The accuracy of the global model declines sharply as the degree of imbalance rises. Throughout the learning phase, the accuracy of the global model under FLPhish($\tau$=0.5) remains at 0.1. It means that FLPhish($\tau$=0.5) considers all clients to be Byzantine, thus invalidating the aggregation process. As can be seen, the FLPhish($\tau$=0.2) outperforms the FLPhish($\tau$=0.1,0.5). To avoid a high false-negative rate, the threshold $\tau$ of FLPhish($\tau$) should be set to a lower value when the distribution becomes non-IID. It should be set to 0.2 in the MNIST experiment. Meanwhile, FLPhish(weight) maintains its performance under varying degrees of imbalance.

    \subsubsection{\ul{Performance Comparison with Different Fractions of Byzantine Clients}} Different fractions of Byzantine clients are taken into account as well. Figure. \ref{fig_table_byzantines_untargeted} shows that the accuracy for FedAvg, Median and Trimmed Mean begins to fall rapidly when the Byzantine portion reaches nearly 50\%. Furthermore, FedAvg, Median, and Trimmed Mean perform extremely invalidly (the accuracy falls below 1\%) when encountered with high fractions of Byzantine attackers. In comparison, FLPhish($\tau$) and FLPhish(weight)'s performance under various thresholds maintain a high level of performance as the portion grows. Both FLPhish($\tau$) and FLPhish(weight) effectively detect Byzantine clients and accurately discard them from the aggregation process. The global model can be successfully trained without the involvement of Byzantine clients during the aggregation procedures. Furthermore, the data show that FLPhish($\tau$=0.1) performs worse than the other two. This is since the threshold=0.1 is far too low to adequately detect Byzantine clients. Though the Byzantine attackers try to make the right predictions of the public dataset and then send the opposite wrong predictions to the FL server. But they also make wrong predictions first and coincidentally transfer the opposite right predictions to the FL server. Thus the reputation values of some Byzantine clients, in particular, can exceed 0.1. It indicates that if the threshold is not well set, Byzantine clients can avoid being detected by FLPhish($\tau$). As the fractions of Byzantine clients go over the threshold of 50\%, the harm brought by untargeted Byzantine attacks increase rapidly for they outnumber the fractions of the benign clients. In the meantime, FLPhish(weight) shows stable performance whenever the fractions of Byzantine clients are.

    \subsection{{Experiment Result Analysis}} Our experiment results demonstrate that both FLPhish($\tau$) and FLPhish(weight) outperform FedAvg, Median \cite{ref_13_defense} and Trimmed Mean \cite{ref_13_defense} on defending against Byzantine attacks in FL. The results also show that the threshold setup has a significant impact on the performance of the FLPhish($\tau$). Only when the managers of FL servers have complete knowledge of FL clients (the imbalance degree of each client), can they accurately set the value of the threshold suitably. However, it is not feasible due to the privacy concern of FL. FLPhish(weight), on the other hand, remains constant regardless of the circumstances. FLPhish(weight) can let clients with higher performance have a stronger influence on the aggregation process by using the reputation value as the weight. To summarize, FLPhish(weight) is a better and more stable aggregation technique than FLPhish(weight) and demonstrates to be a more practical Byzantine defense framework.

\section{Conclusion}
In this paper, we have designed an FL architecture, Ensemble Federated Learning (called EFL) in this study, which allows knowledge transferring between the FL server and FL clients via the unlabeled dataset and FL clients' predictions of it. We have crafted the FLPhish technique to make EFL resistant to Byzantine attacks by using a labeled dataset as `bait' to detect malicious Byzantine clients. Furthermore, we have proposed a reputation technique based on Bayesian inference to determine a client's level of trust. We have proved the security of FLPhish scheme against backdoor attacks. At last, we have tested our suggested FLPhish in a variety of scenarios. The results of the experiment demonstrate the comparable performance of FLPhish in terms of accuracy and robustness under Byzantine attacks in FL. 
\par Our future work will focus on evaluating FLPhish against more advanced Byzantine attacks and improving FLPhish scheme's efficiency.
